---
title: "Gemini paper - Demographics and further analyses"
output:
  word_document:
    number_sections: true
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(binom) # needed for stats
library(ggplot2)
library(ggpubr)
theme_set(theme_pubr())
```


This script calculates numbers excluded from the analysis per category, demographic information, detailed information on the cancers detected and on the additionally arbitrated cases.

# Load data and define AI operating points

```{r load-data}
data <- readRDS("evaluation_dataset.rds")

## stats functions
source("calc-performance-functions.R")

## 4 operating points
OP1 <- 0.557
OP2 <- 0.4829
OP3 <- 0.3957
OP4 <- 0.2248

OPs <- c(OP1, OP2, OP3, OP4)
```

# Exclusions

## Full dataset minus opt-outs

```{r}
df <- data

nrow(df)
```

## Minus technical recalls

```{r}
# remove tech recalls for individual reader opinions
# remove tech recalls for final screen decision
df <- df %>% 
  filter(Reader.1.opinion != "Technical Recall",
         Reader.2.opinion != "Technical Recall",
         Reader.3.opinion != "Technical Recall",
         SBSS.Final.Decision != "TECH_RECALL" | is.na(SBSS.Final.Decision))

nrow(df)
  
```

### Comparison between cases affected and not affected by the IT error - cancer detection rate 

```{r}
# DR: double reading
print("Cancer detection rate for the unaffected cases [per thousand]")
sum(df$cancer_DR[df$potential_read_pre_fix == 1])/(sum(df$potential_read_pre_fix))*1000

print("Cancer detection rate for the affected cases [per thousand]")
sum(df$cancer_DR[df$potential_read_pre_fix == 0])/(sum(df$potential_read_pre_fix == 0))*1000

chisq.test(table(df$cancer_DR, df$potential_read_pre_fix))

```

### Comparison between cases affected and not affected by the IT error - recall rate

```{r}
# RR: recall rate
df_RR <- df %>% 
  mutate(DR = case_when(Overall.opinion == "Review Required" ~ 1, 
                        Overall.opinion == "Routine Recall" ~ 0,
                        TRUE ~ NA_real_)) %>% 
  filter(Overall.opinion == "Review Required" | Overall.opinion == "Routine Recall")

# potential_read_pre_fix: cases which had the potential to be additionally AI
##  read (i.e. cases affected by the IT error, 0: yes, 1: no)
print("Recall rate for the unaffected cases [%]")
sum(df_RR$DR[df_RR$potential_read_pre_fix == 1])/(sum(df_RR$potential_read_pre_fix))*100

print("Recall rate for the affected cases [%]")
sum(df_RR$DR[df_RR$potential_read_pre_fix == 0])/(sum(df_RR$potential_read_pre_fix == 0))*100

chisq.test(table(df_RR$DR, df_RR$potential_read_pre_fix))

```

### Mean age for cases affected and not affected by the IT error

```{r}

print("Mean age for the unaffected cases [%]")
median(df$Age[df$potential_read_pre_fix == 1])

print("Recall rate for the affected cases [%]")
median(df$Age[df$potential_read_pre_fix == 0])

```

## Minus IT error issue

```{r}
df <- df %>% 
  filter(potential_read_pre_fix == 1)

nrow(df)

```


### Comparison between cases read and not read by the AI - cancer detection rate 

```{r, eval = FALSE}

# Mia_read_pre_fix: read by the AI prior to the IT error fix
print("Cancer detection rate for the read cases [per thousand]")
sum(df$cancer_DR[df$Mia_read_pre_fix == 1])/(sum(df$Mia_read_pre_fix))*1000

print("Cancer detection rate for the cases not read [per thousand]")
sum(df$cancer_DR[df$Mia_read_pre_fix == 0])/(sum(df$Mia_read_pre_fix == 0))*1000

chisq.test(table(df$cancer_DR, df$Mia_read_pre_fix))

```


## Minus AI exclusions

```{r}
df <- df %>% 
  filter(Mia_read_pre_fix == 1)

nrow(df)
```

## Minus Review Symptoms

```{r}
# remove if reader 1, 2 or 3 have a Review (Symptoms) opinion
df <- df %>% 
  filter(Reader.1.opinion != "Review (Symptoms)",
         Reader.2.opinion != "Review (Symptoms)",
         Reader.3.opinion != "Review (Symptoms)")

nrow(df)
```

# Demographics

## Age

```{r}
# categorise age and create a summary table
df <- df %>% 
  mutate(age_cat = cut(Age,
                       breaks = c(50, 55, 60, 65, 70, 90), right = FALSE))

table(df$age_cat)
```

## Special Requirements

```{r}

# spec_req: special requirements of a woman attending for breast screening
df$spec_req <- gsub(',', '', df$Special.requirement.list)

# N_req: number of special requirements
df <- df %>%
  mutate(N_req = nchar(spec_req))# %>% 

table(df$N_req)

# create dataframe of cases with exactly one special requirement
df_req_1 <- df %>% 
  filter(Special.requirement == "Yes",
         N_req == 1)

table(df_req_1$Special.requirement.list)
```

## Screen-detected breast cancers

### Size

```{r}
# dataframe which only contains cases diagnosed with a cancer through routine
##  double reading
df_cancer <- filter(df, cancer_DR == 1)

table(df_cancer$cancer_size_cat, useNA = "ifany")
```

### invasive/DCIS

```{r}
table(df_cancer$cancer_type)
```


### ER status

```{r}
# dataframe which only contains invasive cancer cases (receptor status is not
##  assessed for non-invasive cases)
df_receptor <- df_cancer %>% 
  filter(cancer_type == "invasive")

table(df_receptor$ER.Status, useNA = "ifany")
```

### PR status

```{r}
table(df_receptor$PgR.Status, useNA = "ifany")
```

### HER2 status

```{r}
table(df_receptor$HER2.Status, useNA = "ifany")
```

### Triple negative

```{r}
# triple negative cancer: PgR, ER and HER2 are negative
sum(df_receptor$PgR.Status == "Negative" & df_receptor$ER.Status == "Negative" & df_receptor$HER2.Status == "Negative", na.rm = TRUE)
```

# AI standalone recall rate [%] at OP2

```{r Mia-standalone}
# convert raw AI score to recall opinion (1 is yes; 0 is no)
df$Mia_OP2 <- ifelse(df$Mia.Casewise.Raw.Score >= OP2, 1, 0)

round(sum(df$Mia_OP2 == 1)/nrow(df)*100, 1)
```


# AI Additional-Read detected cancers

There are `r sum(df$cancer_XR == 1)` cancers detected with the support of AI.

## Size

```{r}
# XR: additionally (eXtra) Read
# dataframe which only contains XR cancer cases
df_XR_cancer <- filter(df, cancer_XR == 1)
table(df_XR_cancer$XR_size, useNA = "ifany")
```

## Invasive

```{r}
# dataframe which only contains invasive XR cancer cases (receptor status is not
##  assessed for non-invasive cases)
table(df_XR_cancer$Invasive == "y")
```


## ER status

```{r}
df_XR_receptor <- df_XR_cancer %>% 
  filter(Invasive == "y")

table(df_XR_receptor$XR_ER, useNA = "ifany")
```

## PR status

```{r}
table(df_XR_receptor$XR_PR, useNA = "ifany")
```

## HER2 status

```{r}
table(df_XR_receptor$XR_HER2, useNA = "ifany")
```

## Triple negative

```{r}
sum(df_XR_receptor$XR_PR == "negative" & df_XR_receptor$XR_ER == "negative" & df_XR_receptor$XR_HER2 == "Negative", na.rm = TRUE)
```

# Additionally arbitrated cases

## Arbitration time

```{r}
# group the numbers together as some were recorded with a range (e.g. 60-120),
##  while other cases had an exact number recorded

# Discordance: cases eligible for the live AI-Additional Read workflow, i.e.
##  not recalled by routine double reading but flagged by the AI at OP2

df_time <- df %>% 
  mutate(XR_time = recode(df$arbitration.discussion.time..secs.,
                          "0-30" = "15",
                          "30-60" = "45",
                          "60-120" = "90",
                          "120-300" = "210")) %>% 
  mutate(XR_time = as.integer(XR_time)) %>% 
  mutate(XR_time_cat = cut(XR_time,
                       breaks = c(0, 30, 60, 120, 300))) %>% 
  filter(Discordance == "positive")

```

### Categorised by recall decision:

```{r}
table(df_time$XR_time_cat, df_time$XR_recall, useNA = "ifany")
test <- table(df_time$XR_time_cat, df_time$XR_recall, useNA = "ifany")

prop.trend.test(test[,1], test[,1]+test[,2])
```

### Categorised by cancer diagnosis:

```{r}

table(df_time$XR_time_cat, df_time$cancer, useNA = "ifany")

```


## Reason for no recall following additional review

```{r}
# recode duplicate (similar but not the same) coding

df <- df %>% 
  mutate(Reason.for.no.recall...primary = recode(df$Reason.for.no.recall...primary,
                                                 "Low-lying LN" = "low-lying LN",
                                                 "No lesion seen" = "no lesions seen",
                                                 "Previous biopsy" = "previous biopsy",
                                                 "Previous screening film" = "previous screening film",
                                                 "Previous surgery" = "previous surgery",
                                                 "Previous symptomatic films" = "previous symptomatic",
                                                 "Vascular" = "vascular",
                                                 "Well defined" = "well-defined"))

table(df$Reason.for.no.recall...primary)
```

## Additional AI-detected cancers

### Grade
```{r}
# grade_comb: cancer grade combined
# remainder are DCIS
df <- df %>% 
  mutate(Grade_comb = case_when(
    Grade1 == 2 ~ "Grade 2",
    Grade1 == 3 ~ "Grade 3",
    cancer_XR == 1 ~ "DCIS"
  ))

# Grade of tumour
table(df$Grade_comb)

```
### Grade of DCIS

```{r}

# create dataframe of XR non-invasive cancer cases (ductal carcinoma in situ: DCIS)
# convert "intermediate to high" grade to avoid small numbers in output which may impact patient confidentiality

df_XR_DCIS <- df %>% 
  filter(Cancer.Yes.No == "y" | Cancer.Yes.No == "Y",
         Invasive != "y") %>% 
  mutate(Grade = case_when(
    Grade == "intermediate to high" ~ "High",
    TRUE ~ Grade
  ))

table(df_XR_DCIS$Grade)

```

### How many additionally detected cancers had already gone through arbitration?

```{r}

# create dataframe of additionally detected cancer cases that had already gone through regular arbitration

df_XR_arbitration <- df %>% 
  filter(!is.na(Grade_comb),
         Reader.3.opinion != "")
  
nrow(df_XR_arbitration)

```
